### 2018-04-18 concepts.tex
new text, acks

Delta-debugging \cite{DD}\footnote{We use delta-debugging to stand for
  all test reduction algorithms, even those \cite{CReduce,tstl} that do not use the exact
  binary-search that distinguishes delta-debugging \emph{per se}; the
  differences are immaterial for our purposes.}  is a widely used method for reducing the
size of failing tests, making them easier to understand and debug.
Delta-debugging in the context of detecting nondeterminism has two
purposes.  One is simply the usual goal of reducing the size of a
test.  Identifying the cause of nondeterminism may be very easy in a
test consisting of ten library function calls (it is one of these ten
calls), but very difficult in a test consisting of a hundred library
function calls.  This is no different than the common use of
delta-debugging.  However, in horizontal nondeterminism detection,
delta-debugging also tends to \emph{change the probability of
  nondeterministic behavior}; this can be both harmful and
beneficial.  The behavior is part of a more general (and, to our
knowledge, not previously investigate) issue:  using delta-debugging
to minimize a test with respect to a predicate that only holds with a
certain probability (the predicate itself is nondeterministic).  Note
that while most descriptions of delta-debugging discuss reducing a
test with respect to \emph{test failure} (hence the term
``delta-\emph{debugging}'') we adopt the wider view of delta-debugging
as reducing the length of a test while maintaining that an arbitrary
predicate remains true of the shorter versions of the test.  The
traditional predicate is ``the test fails'' but other predicates,
especially those related to code-coverage, can also be useful
\cite{icst2014,stvrcausereduce,NonAdeq}.  Our contribution here is to consider
what happens when the predicate to be evaluated is not deterministic.
``The test behaves flakily'' is a simple, important, example of such a predicate.

@@

The probability that this test behaves nondeterministically is
0.84645, the result of the simple calculation $(1-0.01) \times
(1-0.05) \times (1-0.10)$.  Figure \ref{fig:p1} shows one run of
delta-debugging on a test of length 500.  The initial test is
esentially guaranteed to behave nondeterministically; delta-debugging
only checks that it \emph{can} behave nondeterministically, and so, in
the course of reducing the test length to a test with only two
operations, the delta-debugging algorithm also reduces the probability
of nondeterminism to only about 20\%.  If we use a predicate that
``forces'' the test to behave nondeterministically at least half the
time, by sampling the predicate value 100 times and only returning
true when at least 50 of the evaluations report true, we see the
behavior in Figure \ref{fig:p2}:  the final test is slightly longer,
but still falls well short of our target of exhibiting nondeterminism
50\% of the time.  Finally, Figure \ref{fig:p3} shows what happens if
we use the same target of 50\% nondeterminism, but use only 10
samples, with 10 replications ($N$ = 10, $M$ = 10):  the test is not
much longer than in Figure \ref{fig:p2}, but the probability of
nondeterminism is above our target value, close to 60\%.

>>>>

Delta-debugging \cite{DD}\footnote{We use delta-debugging to stand for
  all test reduction algorithms, even those \cite{CReduce,tstl} that do not use the exact
  binary-search that distinguishes delta-debugging \emph{per se}; the
  differences are immaterial for our purposes.}  is a widely used method for reducing the
size of failing tests, making them easier to understand and debug.
Delta-debugging in the context of detecting nondeterminism has two
purposes.  One is simply the usual goal of reducing the size of a
test.  Identifying the cause of nondeterminism may be very easy in a
test consisting of ten library function calls (it is one of these ten
calls), but very difficult in a test consisting of a hundred library
function calls.  This is no different than the common use of
delta-debugging.  However, in horizontal nondeterminism detection,
delta-debugging also tends to \emph{change the probability of
  nondeterministic behavior}; this can be both harmful and
beneficial.  The behavior is part of a more general (and, to our
knowledge, not previously investigated) issue:  using delta-debugging
to minimize a test with respect to a predicate that only holds with a
certain probability (the predicate itself is nondeterministic).  Note
that while most descriptions of delta-debugging discuss reducing a
test with respect to \emph{test failure} (hence the term
``delta-\emph{debugging}'') we adopt the wider view of delta-debugging
as reducing the length of a test while maintaining that an arbitrary
predicate remains true of the shorter versions of the test.  The
traditional predicate is ``the test fails'' but other predicates,
especially those related to code-coverage, can also be useful
\cite{icst2014,stvrcausereduce,NonAdeq}.  Our contribution here is to consider
what happens when the predicate to be evaluated is not deterministic.
``The test behaves flakily'' is a simple, important, example of such a predicate.

@@

The probability that this test behaves nondeterministically is
0.84645, the result of the simple calculation $(1-0.01) \times
(1-0.05) \times (1-0.10)$.  Figure \ref{fig:p1} shows one run of
delta-debugging on a test of length 500.  The initial test is
esentially guaranteed to behave nondeterministically; delta-debugging
only checks that it \emph{can} behave nondeterministically, and so, in
the course of reducing the test length to a test with only two
operations, the delta-debugging algorithm also reduces the probability
of nondeterminism to only about 20\%.  If we use a predicate that
``forces'' the test to behave nondeterministically at least half the
time, by sampling the predicate value 100 times and only returning
true when at least 50 of the evaluations report true, we see the
behavior in Figure \ref{fig:p2}:  the final test is slightly longer,
but still falls well short of our target of exhibiting nondeterminism
50\% of the time.  Finally, Figure \ref{fig:p3} shows what happens if
we use the same target of 50\% nondeterminism, but use only 10
samples, with 10 replications ($N$ = 10, $M$ = 10):  the test is not
much longer than in Figure \ref{fig:p2}, but the probability of
nondeterminism is above our target value, close to 60\%.  Note that
this example is purely monotonic; the pattern of probability changes
in non-monotonic settings can be even more difficult to predict and
control, but the principle of forcing a probability by sampling and
replication still holds.

### 2018-04-18 conclusions.tex
new text, acks

>>>>

More generally, we are interested in using test decomposition
\cite{Composition} to more easily isolate, understand, and avoid
nondeterminism is tests, and to mitigate the problem of flaky tests.
Reliable detection of nondeterminism/flakiness is the first step to
active solutions to eliminating nondeterministic/flaky test behavior.

{\scriptsize {\bf Acknowledgements:}  The authors would like to thank John Micco,
Jeff Listfield, and Celal Ziftci, all at Google, for discussions of
the problem of flaky tests, Andreas Zeller and David R. MacIver for discussion of the
problem of delta-debugging of nondeterministic properties, and Chris
Colohan for discussion of sources of process-based nondeterminism.}

### 2018-04-18 implementation.tex
new text, acks

We therefore instead made nondeterminism detection a first-class
property in TSTL, using TSTL's existing notation for marking some
types of values as \emph{opaque} (not usefully compared for equality),
used in automatic abstraction-based testing.  Tables \ref{tab:methods}
and \ref{tab:options} show, respectively, the primary additions to the
TSTL Python API and the TSTL command line tools.  The TSTL test
generation tool is called {\tt rt}, since the default mode is a fast
pure Random Tester, and {\tt reduce} is TSTL's tool for
delta-debugging and test
normalization (more aggressive reduction that also tends to ease
debugging and test triage by making tests failing due to the same
fault more similar \cite{onetest}).  With these additions,
developers of TSTL-based testing tools, or developers using TSTL tools
to test code can easily add nondeterminism to the set of properties
checked.

>>>>

We therefore instead made nondeterminism detection a first-class
property in TSTL, using TSTL's existing notation for marking some
types of values as \emph{opaque} (not usefully compared for equality),
used in automatic abstraction-based testing.  In TSTL, a test is a
sequence of \emph{actions}; parameter values to method and function
calls of the Software Under Test (SUT) are stored in finite-sized variable \emph{pools}.  A
TSTL test is (essentially) a sequence of assignments to pool values,
and method/function calls using pool values (including method calls on
pool objects).  This formalism is a widely used and efficient way to
represent unit tests \cite{Pachecho,AndrewsTR}.  In TSTL, \emph{visible
value determinism} is based on comparing the values of \emph{all pool
variables} after each test action (pool values other than the one
assigned to must also be compared, since a common source of
nondeterminism is when a call results in a change to a value passed as
a parameter, or even due to a shared reference held by two values).
TSTL omits comparison only of pools marked as {\tt OPAQUE} in the
TSTL language test harness (and values that do not support equality
checking).  \emph{Final State Determinism} simply performs the same
comparison, but only on the final values of all pool variables after a test has
finished executing.

 Tables \ref{tab:methods}
and \ref{tab:options} show, respectively, the primary additions to the
TSTL Python API and the TSTL command line tools.  The TSTL test
generation tool is called {\tt rt}, since the default mode is a fast
pure Random Tester, and {\tt reduce} is TSTL's tool for
delta-debugging and test
normalization (more aggressive reduction that also tends to ease
debugging and test triage by making tests failing due to the same
fault more similar \cite{onetest}).  With these additions,
developers of TSTL-based testing tools, or developers using TSTL tools
to test code can easily add nondeterminism to the set of properties
checked.
